Hardware-accelerated Rendering of Intersecting Volumes using Boolean Operations

### 2013-04-22 paper.tex
Replaced XML for second approach with pseudocode

\begin{Verbatim}[fontsize=\small]
  <texture id="first-volume-texture" file="first.vlb">
  <texture id="second-volume-texture" file="second.vlb">
  <texture id="coords-texture-1" width="768" height="768">
  <texture id="coords-texture-2" width="768" height="768">
  <texture id="results-texture" width="768" height="768">
\end{Verbatim}

Next the part of the volume behind the intersection is rendered into the results
texture.  To get just that part the back faces of the Boolean AND are used
instead of the front faces of the cube.  Note in the code snippet below we omit
the instructions for rendering the second volume and for setting uniform
variables.  We also put the Boolean AND into a group so it can be redrawn later
without duplicating the calculations or geometry.

\begin{Verbatim}[fontsize=\small]
  <cull faces="front" />
  <framebuffer>
    <attachment usage="color" link="coords-texture-1" />
    <use program="primary-coords-program">
      <instance of="cube-1-group" />
    </use>
  </framebuffer>
  <framebuffer>
    <attachment usage="color" link="results-texture" />
    <use program="first-pass-program">
      <group id="intersection-group">
        <booleanAnd of="cube-1 cube-2" />
      </group>
    </use>
  </framebuffer>
\end{Verbatim}

@@

\begin{Verbatim}[fontsize=\small]
  <cull faces="front" />
  <framebuffer>
    <attachment usage="color" link="coords-texture-1" />
    <use program="primary-coords-program">
      <instance of="intersection-group" />
    </use>
  </framebuffer>
  <framebuffer>
    <attachment usage="color" link="coords-texture-2" />
    <use program="secondary-coords-program">
      <instance of="intersection-group" />
    </use>
  </framebuffer>
  <cull faces="back" />
  <framebuffer>
    <attachment usage="color" link="third-pass-texture" />
    <use program="third-pass-program">
      <uniform name="CoordsTexture1" link="coords-texture-1" />
      <uniform name="CoordsTexture2" link="coords-texture-2" />
      <instance of="intersection-group" />
    </use>
  </framebuffer>
\end{Verbatim}

After that the part of the volume in front of the intersection is rendered to
the results texture.  We have the shader program set the depth of the fragments
generated to the lowest value to effectively mask off that part of the screen.

\begin{Verbatim}[fontsize=\small]
  <cull mode="back" />
  <framebuffer>
    <attachment usage="color" link="coords-texture-1" />
    <use program="primary-coords-program">
      <uniform name="MVPMatrix" usage="modelviewprojection" />
      <instance of="cube-1-group" />
    </use>
  </framebuffer>
  <framebuffer>
    <attachment usage="color" link="results-texture" />
    <use program="fourth-pass-program">
      <instance of="intersection-group" />
    </use>
  </framebuffer>
\end{Verbatim}

@@

\begin{Verbatim}[fontsize=\small]
  <texture id="volume-texture-1" file="volume-1.vlb" />
  <texture id="volume-texture-2" file="volume-2.vlb" />
  <texture id="coords-texture-1" size="768" />
  <texture id="coords-texture-2" size="768" />
  <texture id="results-texture" size="768" />
\end{Verbatim}

@@

\begin{Verbatim}[fontsize=\small]
  <renderbuffer id="depth-renderbuffer" format="depth" size="768" />
\end{Verbatim}

@@

\begin{Verbatim}[fontsize=\small]
  <framebuffer>
    <attachment usage="color" link="results-texture" />
    <clear color="0 0 0 0" />
  </framebuffer>
\end{Verbatim}

Then the pieces behind the intersection should be used to render the volumes to
the results texture.  Because there may be multiple pieces behind the
intersection and we only want the texture coordinates of the faces that are
farthest back, we first clear the depth buffer to zero and set the depth
function to \emph{GREATER}.  Once we have those coordinates stored, we do the
opposite by clearing the depth buffer to one and setting the depth function back
to \emph{LESS}.  That way we render using the faces that are farthest forward.
Note in the code snippet below we only show the first volume.

\begin{Verbatim}[fontsize=\small]
  <framebuffer>
    <attachment usage="color" link="coords-texture-1" />
    <attachment usage="depth" link="depth-renderbuffer" />
    <clear depth="0.0" />
    <depth function="greater" />
    <cull faces="front" />
    <use program="primary-coords-program">
      <group id="pieces-of-cube-1-behind-intersection-group">
        <booleanXor of="cube-1 cube-2" hide="cube-2" filter="gequal" />
      </group>
    </use>
  </framebuffer>
  <framebuffer>
    <attachment usage="color" link="results-texture" />
    <attachment usage="depth" link="depth-renderbuffer" />
    <clear depth="1.0" />
    <depth function="less" />
    <cull faces="back" />
    <use program="first-pass-program">
      <uniform name="MVPMatrix" usage="viewprojection" />
      <uniform name="BackFacesTexture" link="coords-texture-1" />
      <uniform name="VolumeTexture" link="volume-texture-1" />
      <instance of="pieces-of-cube-1-behind-intersection-group" />
    </use>
  </framebuffer>
\end{Verbatim}

@@

\begin{Verbatim}[fontsize=\small]
  <cull faces="front" />
  <framebuffer>
    <attachment usage="color" link="coords-texture-1" />
    <use program="primary-coords-program">
      <uniform name="MVPMatrix" usage="viewprojection" />
      <group id="intersection-group">
        <booleanAnd of="cube-1 cube-2" />
      </group>
    </use>
  </framebuffer>
  <framebuffer>
    <attachment usage="color" link="coords-texture-2" />
    <use program="secondary-coords-program">
      <uniform name="MVPMatrix" usage="viewprojection" />
      <instance of="intersection-group" />
    </use>
  </framebuffer>
  <cull faces="back" />
  <framebuffer>
    <attachment usage="color" link="results-texture" />
    <use program="second-pass-program">
      <uniform name="MVPMatrix" usage="viewprojection" />
      <uniform name="ResultsTexture" link="results-texture" />
      <uniform name="CoordsTexture1" link="coords-texture-1" />
      <uniform name="CoordsTexture2" link="coords-texture-2" />
      <uniofrm name="VolumeTexture1" link="volume-texture-1" />
      <uniofrm name="VolumeTexture2" link="volume-texture-2" />
      <instance of="intersection-group" />
    </use>
  </framebuffer>
\end{Verbatim}

Next we render any pieces of the first cube in front of the intersection to the
results texture.  Similar to the first pass, we use the depth buffer to make
sure we only take the texture coordinates of the faces we want.  Again we omit
the code to render the second cube.

\begin{Verbatim}[fontsize=\small]
  <framebuffer>
    <attachment usage="color" link="coords-texture-1" />
    <attachment usage="depth" link="depth-renderbuffer" />
    <cull faces="front" />
    <depth function="greater" />
    <clear depth="0.0" />
    <use program="primary-coords-program">
      <uniform name="MVPMatrix" usage="viewprojection" />
      <group id="pieces-of-cube-1-in-front-of-intersection">
        <booleanXor of="cube-1 cube-2" hide="cube-2" filter="less" />
      </group>
    </use>
  </framebuffer>
  <framebuffer>
    <attachment usage="color" link="results-texture" />
    <attachment usage="depth" link="depth-renderbuffer" />
    <depth function="less" />
    <cull faces="back" />
    <use program="third-pass-program">
      <uniform name="MVPMatrix" usage="viewprojection" />
      <uniform name="ResultsTexture" link="results-texture" />
      <uniform name="CoordsTexture" link="coords-texture-1" />
      <uniform name="VolumeTexture" link="volume-texture-1" />
      <insance of="pieces-of-cube-1-in-front-of-intersection" />
    </use>
  </framebuffer>
\end{Verbatim}

@@

\begin{Verbatim}[fontsize=\small]
  <use program="final-pass-program">
    <uniform name="ResultsTexture" link="results-texture" />
    <translate x="-1.0">
      <uniform name="MVPMatrix" usage="modelviewprojection" />
      <cube id="cube-1" />
    </translate>
    <translate x="+1.0">
      <uniform name="MVPMatrix" usage="modelviewprojection" />
      <cube id="cube-2" />
    </translate>
  </use>
\end{Verbatim}

>>>>

\begin{enumerate}
  \item Bind volume textures
  \item Bind textures to store texture coordinates
  \item Bind texture to store results
\end{enumerate}

Next the part of the volume behind the intersection is rendered into the results
texture.  To get just that part the back faces of the Boolean AND are used
instead of the front faces of the cube.

\begin{enumerate}
  \item Cull front faces
  \item Bind framebuffer with coordinates texture
  \item Activate shader program for rendering texture coordinates
  \item Draw cube behind intersection
  \item Bind framebuffer with results texture
  \item Activate shader program for rendering a volume
  \item Draw Boolean AND of cubes
\end{enumerate}

@@

\begin{enumerate}
  \item Cull front faces
  \item Bind framebuffer with first coordinates texture
  \item Activate shader program for rendering primary texture coordinates
  \item Draw Boolean AND of cubes
  \item Bind framebuffer with second coordinates texture
  \item Activate shader program for rendering secondary texture coordinates
  \item Draw Boolean AND of cubes
  \item Cull back faces
  \item Bind framebuffer with results texture
  \item Activate shader program for rendering volume
  \item Draw Boolean AND of cubes
\end{enumerate}

After that the part of the volume in front of the intersection is rendered to
the results texture.

\begin{enumerate}
  \item Cull back faces
  \item Bind framebuffer with coordinates texture
  \item Activate shader program for rendering primary texture coordinates
  \item Draw cube in front of intersection
  \item Bind framebuffer with results texture
  \item Activate shader program for rendering volume
  \item Draw Boolean AND of cubes
  \item Mask off depth buffer
\end{enumerate}

The area of the screen with the intersection should be masked off so that
nothing else in that area will be rendered.

\begin{enumerate}
  \item Activate shader program for setting depth to zero
  \item Draw Boolean AND of cubes
\end{enumerate}

@@

\begin{enumerate}
  \item Bind volume textures
  \item Bind textures to store texture coordinates
  \item Bind texture to store results
\end{enumerate}

@@

\begin{enumerate}
  \item Create depth renderbuffer
\end{enumerate}

@@

\begin{enumerate}
  \item Bind framebuffer with results texture
  \item Clear color
\end{enumerate}

Then the pieces behind the intersection should be used to render the volumes to
the results texture.  Because there may be multiple pieces behind the
intersection and we only want the texture coordinates of the faces that are
farthest back, we first clear the depth buffer to zero and set the depth
function to \emph{GREATER}.  Once we have those coordinates stored, we do the
opposite by clearing the depth buffer to one and setting the depth function back
to \emph{LESS}.  That way we render using the faces that are farthest forward.

\begin{enumerate}
  \item Bind framebuffer with coordinates texture and depth renderbuffer
  \item Clear depth to zero
  \item Set depth function to \emph{GREATER}
  \item Cull front faces
  \item Activate shader program for rendering primary texture coordinates
  \item Draw Boolean XOR with pieces behind intersection
  \item Bind framebuffer with results texture and depth renderbuffer
  \item Clear depth to one
  \item Set depth function to \emph{LESS}
  \item Cull back faces
  \item Activate shader program for rendering volumes
  \item Draw Boolean XOR with pieces behind intersection
\end{enumerate}

@@

\begin{enumerate}
  \item Cull front faces
  \item Bind framebuffer with first coordinates texture
  \item Activate shader program for rendering primary texture coordinates
  \item Draw Boolean AND of cubes
  \item Bind framebuffer with second coordinates texture
  \item Activate shader program for rendering secondary texture coordinates
  \item Draw Boolean AND of cubes
  \item Cull back faces
  \item Bind framebuffer with results texture
  \item Activate shader program for rendering volumes
  \item Draw Boolean AND of cubes
\end{enumerate}

Next we render any pieces of the cube in front of the intersection to the
results texture.  Similar to the first pass, we use the depth buffer to make
sure we only take the texture coordinates of the faces we want.

\begin{enumerate}
  \item Bind framebuffer with coordinates texture and depth renderbuffer
  \item Cull front faces
  \item Set depth function to \emph{GREATER}
  \item Clear depth buffer to zero
  \item Activate shader program for rendering primary texture coordinates
  \item Draw Boolean XOR with pieces in front of intersection
  \item Bind framebuffer with results texture and depth renderbuffer
  \item Cull back faces
  \item Set depth function to \emph{LESS}
  \item Activate shader program for rendering volumes
  \item Draw Boolean XOR with pieces in front of intersection
\end{enumerate}

@@

\begin{enumerate}
  \item Activate shader for rendering results
  \item Render both cubes
\end{enumerate}
